---
title: Caret simple modelling - COX2 example
date: 2021-02-18
output: 
  html_document:
    theme: spacelab
    highlight: haddock
    toc: true
    toc_float: true
    toc_depth: 2
    df_print: paged
    number_sections: true
    
author: Fernando Carazo^1^ 
# bibliography: "references.bib" # Create file and uncomment. Cite as ej [@Ignatiadis_2016]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library("knitcitations")
#cleanbib()
options("citation_format" = "pandoc")
# write.bibtex(file="references.bib")
```

Contact info: **Fernando Carazo** <fcarazo@tecnun.es>

[*^1^Department of Biomedical Engineering and Sciences, Tecnun. University of Navarre*](https://www.unav.edu/en/web/departamento-de-ingenieria-biomedica-y-ciencias/investigacion/computational-biology/presentacion)



# COX-2 Activity Data
From Sutherland, O’Brien, and Weaver (2003): A set of 467 cyclooxygenase-2 (COX-2) inhibitors has been assembled from the published work of a single research group, with in vitro activities against human recombinant enzyme expressed as IC50 values ranging from 1 nM to >100 uM (53 compounds have indeterminate IC50 values).

A set of 255 descriptors (MOE2D and QikProp) were generated. To classify the data, we used a cutoff of $2^{2.5}$ to determine activity.

Using data(cox2) exposes three R objects: `cox2Descr` is a data frame with the descriptor data, `cox2IC50` is a numeric vector of IC50 assay values and `cox2Class` is a factor vector with the activity results.

```{r}
library(recipes)
library(caret)
library(vip)
data(cox2)
cox2 <- cox2Descr
cox2$potency <- cox2IC50


# Final transformation
  ## Create the model recipe outcome: "potency", predictors: all variables.

  ## Log10 the outcome

  ## Remove near to zero predictors

  ## find a step to run PCA

  ## Surface area predictors are highly correlated so
  ## conduct PCA just on these.


  ## Remove other highly correlated predictors (up to 0.9 cor)
 

  ## Center and scale all of the non-PCA predictors


```



https://recipes.tidymodels.org/reference/index.html
