---
title: Caret simple modelling - COX2 example
date: 2021-02-18
output: 
  html_document:
    theme: spacelab
    highlight: haddock
    toc: true
    toc_float: true
    toc_depth: 2
    df_print: paged
    number_sections: true
    
author: Fernando Carazo^1^ 
# bibliography: "references.bib" # Create file and uncomment. Cite as ej [@Ignatiadis_2016]
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library("knitcitations")
#cleanbib()
options("citation_format" = "pandoc")
# write.bibtex(file="references.bib")
```

Contact info: **Fernando Carazo** <fcarazo@tecnun.es>

[*^1^Department of Biomedical Engineering and Sciences, Tecnun. University of Navarre*](https://www.unav.edu/en/web/departamento-de-ingenieria-biomedica-y-ciencias/investigacion/computational-biology/presentacion)



# COX-2 Activity Data
From Sutherland, O’Brien, and Weaver (2003): A set of 467 cyclooxygenase-2 (COX-2) inhibitors has been assembled from the published work of a single research group, with in vitro activities against human recombinant enzyme expressed as IC50 values ranging from 1 nM to >100 uM (53 compounds have indeterminate IC50 values).

A set of 255 descriptors (MOE2D and QikProp) were generated. To classify the data, we used a cutoff of $2^{2.5}$ to determine activity.

Using data(cox2) exposes three R objects: `cox2Descr` is a data frame with the descriptor data, `cox2IC50` is a numeric vector of IC50 assay values and `cox2Class` is a factor vector with the activity results.

```{r}
library(recipes)
library(caret)
library(vip)
data(cox2)
cox2 <- cox2Descr
cox2$potency <- cox2IC50


# aux <- cox2[250:256]
# cox2_recipe <-  recipe(potency ~ ., data = aux) %>%
#   ## Log the outcome
#   step_log(potency, base = 10) %>%
#   ## Remove sparse and unbalanced predictors
#   step_nzv(all_predictors()) %>%
#   ## Surface area predictors are highly correlated so
#   ## conduct PCA just on these.
#   step_pca(all_predictors(), threshold = .95, prefix = "surf_area")
# 
# 
# %>%
#   ## Remove other highly correlated predictors
#   step_corr(all_predictors(), -contains("VSA"),
#             threshold = .90) %>%
#   ## Center and scale all of the non-PCA predictors
#   step_center(all_predictors(), -contains("VSA")) %>%
#   step_scale(all_predictors(), -contains("VSA"))
# 
# 
# cox2_transform <- cox2_recipe %>% prep(aux) %>% juice(); cox2_recipe


# Final transformation
  ## Create the model recipe outcome: "potency", predictors: all variables.
cox2_recipe <- recipe(potency ~ ., data = cox2) %>%
  ## Log10 the outcome
  step_log(potency, base = 10) %>%
  ## Remove near to zero predictors
  step_nzv(all_predictors()) %>%
  ## find a step to run PCA
  ## Surface area predictors are highly correlated so
  ## conduct PCA just on these.
  step_pca(contains("VSA"), prefix = "surf_area_",
           threshold = .95) %>%
  ## Remove other highly correlated predictors (up to 0.9 cor)
  step_corr(all_predictors(), -starts_with("surf_area_"),
            threshold = .90) %>%
  ## Center and scale all of the non-PCA predictors
  step_center(all_predictors(), -starts_with("surf_area_")) %>%
  step_scale(all_predictors(), -starts_with("surf_area_"))

set.seed(888)
cox2_lm <- train(cox2_recipe,
                 data = cox2,
                 method = "lm",
                 trControl = trainControl(method = "cv", number = 5))

summary(cox2_lm)
vip(cox2_lm)

```



https://recipes.tidymodels.org/reference/index.html
